---
title: "Aides Drug"
author: "aubrey!"
date: "06/20/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
druggies= select(triUmp, c(treatment, infected, DrugUser))
druggies = filter(druggies, DrugUser == "yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: This vaccine is not effective on drug users.

$H_A$: This vaccine is effective on drug users.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

treatment is a categorical variable

infected is a Categorical variable

so we use the CAT~CAT helper file

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !--> 

```{r}
barchartGC(~infected+ treatment ,data= druggies)
barchartGC(~infected+ treatment ,data= druggies, type="percent")

```
The bar chart shows those who did and didn't get infected among the drug users, based on weather they got the placebo or triUmpsal. Meaning the vaccine is much more effectice than the placebo.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected+ treatment ,data= druggies)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that those who got the drug, the odd ratio of them actually getting AIDS is about a  0.25 : 1 odds.
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows that the p-vaule is 0.0000001242, which is less than 0.05, meaning we jected the null hypothesis. The test also showed that the statitical differnce in those who did and didn't get AIDS based on weather they got placebo or drug, there was only a 27% from those who did get it. Meaning there is not a big statisical difference. 

The fishers exact test shows that the p-vaule is 0.0000000423, which is less than 0.05, so we reject the null hypothesis. It also shows that the odds of getting AIDS is 0.35341285.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Under the impression the statement and question where a new  preventative aides drug (triUmpsal) is under test in clinical trials. Then considering only the Drug Users, does this drug seem effective?  After many test, the P-value for the The chi-square test shows that the p-value is 0.0000001242, which is less than 0.05, meaning we rejected the null hypothesis. The test also showed that the statistical difference in those who did and didn't get AIDS based on weather they got placebo or drug, there was only a 27% from those who did get it. Meaning there is not a big statistical difference. According to The fishers exact test shows that the p-vaule is 0.0000000423, which is less than 0.05, so we reject the null hypothesis. It also shows that the odds of getting AIDS is 0.35341285. In which the results conclude that TriUmpsal is not very effective on druggies in terms of getting infected.
